---
title: "Shiny Gadgets: Cool Helpers in Data Analysis"
subtitle: "A lesser known side of Shiny"
author: "StatistikinDD"
date: "Presentation Created: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["xaringan-themer.css", "libs/_css/my_css.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: 16:9
---

```{r setup, include = FALSE}

options(htmltools.dir.version = FALSE)
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = "")
knitr::opts_chunk$set(echo = TRUE, comment = "")
xaringanthemer::style_mono_accent(title_slide_background_color = "#EDF4FB",
                                  title_slide_text_color = "#000000",
                                  base_font_size = "18px")
Sys.setenv(LANG = "en")

```

# A Working Example: Brush Points

```{r, echo = FALSE}

library(shiny)
library(miniUI)
library(ggplot2)
theme_set(theme_bw(base_size = 16))

ggbrush <- function(data, xvar, yvar) {
  ui <- miniPage(
    gadgetTitleBar("Drag to select points"),
    miniContentPanel(
      # The brush = "brush" argument means we can listen for
      # brush events on the plot using input$brush.
      plotOutput("plot", height = "100%", brush = "brush")
    )
  )
  
  server <- function(input, output, session) {

    # Render the plot
    output$plot <- renderPlot({
      # Plot the data with x/y vars indicated by the caller
      ggplot(data, aes({{xvar}}, {{yvar}})) +
        geom_point()
    })
  
    # Handle the Done button being pressed.
    observeEvent(input$done, {
      # Return the brushed points. See ?shiny::brushedPoints.
      stopApp(brushedPoints(data, input$brush))
    })
  }
  
  runGadget(ui, server)
}

ggbrush(mtcars, hp, mpg)
```

Credit to the RStudio Team: https://shiny.rstudio.com/articles/gadgets.html

---

# Second Example: gglabeller

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# library(devtools)
# install_github("AliciaSchep/gglabeller")

library(gglabeller)
library(ggplot2)
library(ggrepel)

p <- ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point() +
  theme_classic(base_size = 18)

gglabeller_example <- gglabeller(p, aes(label = rownames(mtcars)))

gglabeller_example
```

Credit to Alicia Schep - see gglabeller on Github

---

# R Code: First Example

.pull-left[
```{r ggbrush-code1, echo = TRUE, eval = FALSE}
library(shiny)
library(miniUI)
library(ggplot2)
theme_set(theme_bw(base_size = 16))

ggbrush <- function(data, xvar, yvar) { #<<
  ui <- miniPage( #<<
    gadgetTitleBar("Drag to select points"),
    miniContentPanel(
      # The brush = "brush" argument means we can listen for
      # brush events on the plot using input$brush.
      plotOutput("plot", height = "100%", brush = "brush")
    )
  )
  

```

* Credit to the RStudio Team: https://shiny.rstudio.com/articles/gadgets.html
* Note: I slightly updated the code, using {{}} to pass variables to ggplot()

]

.pull-right[
```{r ggbrush-code2, echo = TRUE, eval = FALSE}
  server <- function(input, output, session) { #<<

    # Render the plot
    output$plot <- renderPlot({ #<<
      # Plot the data with x/y vars indicated by the caller
      ggplot(data, aes({{xvar}}, {{yvar}})) +
        geom_point()
    })
  
    # Handle the Done button being pressed.
    observeEvent(input$done, { #<<
      # Return the brushed points. See ?shiny::brushedPoints.
      stopApp(brushedPoints(data, input$brush))
    })
  }
  
  runGadget(ui, server) #<<
}

ggbrush(mtcars, hp, mpg) #<<
```
]

---

# Shiny Gadgets vs. Shiny Apps

.pull-left[
## Shiny App

.content-box-grey[
### Present output of an analysis
### For end users
### Often hosted on server
]]

--

.pull-right[
## Shiny Gadget

.content-box-gray[
### Use during analysis
### For enhancing programming experience
### For R users
]]

---

# Resources

### gglabeller comes as an R package
* Written by **Alicia Schep**
* install from github: `devtools::install_github("AliciaSchep/gglabeller")`

### Customizations

* Set *viewer* parameter in *runGadget()*: pane / dialog; set width and height 
* You can register Shiny Gadgets as **RStudio Addins**

### Learn more

* RStudio Website: https://shiny.rstudio.com/articles/gadgets.html
* Webinar, **Garrett Grolemund**: https://resources.rstudio.com/webinars/shiny-gadgets
* **Hadley Wickham**'s presentation at Shiny Dev Con 2016: https://resources.rstudio.com/shiny-developer-conference/shinydevcon-gadgets-hadleywickham-1080p?_ga=2.187073576.1466710020.1638966460-723354145.1599642879


---

class: center, middle

# Thanks!

### Youtube: StatistikinDD

### Twitter: @StatistikinDD

### github: fjodor

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr), and [R Markdown](https://rmarkdown.rstudio.com).

Thanks to **Yihui Xie** for *{knitr}* and *{xaringan}* and **Garrick Aden-Buie** for *{xaringanthemer}*.